---
title: "Descriptive Statistics"
output:
  html_document: default
  pdf_document: default
---
# Descriptive Statistics for our Merging Dataset
```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(ggplot2)
library(Hmisc)
library(stringr)
library(plm)
library(knitr)
library(xlsx)
library(dplyr)
library(kableExtra)
options(scipen = 999)

setwd("~/Google Drive/Non-Academic Work/Research/Traina/occupation-productivity-2/")

#starting this year we have files
'%ni%' <- Negate('%in%')

#summary statistics function
summaryStats <- function (x, name, var) {
  print(paste(var, name, sep = " - "))
  stats_vector <- c(mean(x, na.rm = TRUE), sd(x, na.rm = TRUE), mean(is.na(x)))
  names(stats_vector) <- c("Mean", "SD", "Missing")
  return(kable(c(stats_vector, quantile(x, c(0,.1,.25, .50,  .75, .9,1), na.rm = TRUE)),
        col.names = c(var)) %>% kable_classic_2(full_width = F))
}
merge_table_default <- fread("merge_table.csv")
merge_table <- merge_table_default
knitr::opts_chunk$set
```
## Descriptive Statistics for Single Variables
### Occupation Codes
number of unique occupation codes: `r length(unique(merge_table$OCC_CODE))`

distribution of unique occupation codes by # digits:
```{r base_occupation_codes1, echo = FALSE, results='asis', warning = FALSE}
kable(table(unique(merge_table[,c("OCC_CODE", "LEVEL_OCC_OE"), with = FALSE])$LEVEL_OCC_OE), col.names = c("occ digits", "freq"))  %>%
  kable_classic_2(full_width = F)
```
overall distribution of occupation codes in table
```{r base_occupation_codes2, echo = FALSE, results='asis', warning = FALSE}
kable(table(merge_table$LEVEL_OCC_OE), col.names = c("occ digits", "freq")) %>%
  kable_classic_2(full_width = F)
```

### Industry Codes
number of unique industry codes: `r length(unique(merge_table$NAICS))`

distribution of unique industry codes by # digits
```{r base_industry_codes1, echo = FALSE, results='asis'}
kable(table(unique(merge_table[,c("OCC_CODE", "LEVEL_NAICS_OE"), with = FALSE])$LEVEL_NAICS_OE), col.names = c("industry digits", "freq"))  %>%
  kable_classic_2(full_width = F)
```
overall distribution of industry codes in table
```{r base_industry_codes2, echo = FALSE, results='asis'}
kable(table(merge_table$LEVEL_NAICS_OE), col.names = c("industry digits", "freq")) %>%
  kable_classic_2(full_width = F)
```


### Year Codes
number of years,`r length(unique(merge_table$YEAR))`

table of years
```{r base_year_codes, echo = FALSE, results='asis'}
kable(table(merge_table$YEAR), col.names = c("year", "freq")) %>%
  kable_classic_2(full_width = F)
```
### Area Codes
number of area codes, `r length(unique(merge_table$AREA_CODE))`

distribution of a unique set of codes & all codes by area classification 
```{r base_area_codes, echo = FALSE, results='asis'}
temp <- merge_table[,c("AREA_CODE"), with = FALSE]
temp[AREA_CODE == 99, classification := "national"]
temp[is.na(classification) & AREA_CODE<100, classification := "state/territory"]
temp[is.na(classification) & AREA_CODE<100000 & AREA_CODE %% 10 == 0, classification := "city"]
temp[is.na(classification) & AREA_CODE<100000, classification := "metropolitan area"]
temp[is.na(classification), classification := "non metropolitan area"]
kable(table(unique(temp[,c("AREA_CODE", "classification"), with = FALSE])$classification),
      col.names = c("area classification", "freq")) %>%
  kable_classic_2(full_width = F)
kable(table(temp$classification), col.names = c("area classification", "freq")) %>%
  kable_classic_2(full_width = F)
```
table of years
```{r base_area_codes2, echo = FALSE, results='asis'}
kable(table(temp$classification), col.names = c("area classification", "freq")) %>%
  kable_classic_2(full_width = F)
```

## Two Variable Pairs
number of unique industry-year pairs, `r dim(unique(merge_table[,c("NAICS", "YEAR"), with = FALSE]))[1]`

number of unique industry-occupation pairs, `r dim(unique(merge_table[,c("NAICS", "OCC_CODE"), with = FALSE]))[1]`

number of unique industry-area pairs, `r dim(unique(merge_table[,c("NAICS", "AREA_CODE"), with = FALSE]))[1]`

number of unique year-occupation pairs, `r dim(unique(merge_table[,c("YEAR", "OCC_CODE"), with = FALSE]))[1]`

number of unique year-area pairs, `r dim(unique(merge_table[,c("YEAR", "AREA_CODE"), with = FALSE]))[1]`

number of unique occupation-area pairs, `r dim(unique(merge_table[,c("OCC_CODE", "AREA_CODE"), with = FALSE]))[1]`

## Three Variable Pairs
number of unique industry-year-occupation groups, `r dim(unique(merge_table[,c("NAICS", "YEAR", "OCC_CODE"), with = FALSE]))[1]`

number of unique industry-year-area groups, `r dim(unique(merge_table[,c("NAICS", "YEAR", "AREA_CODE"), with = FALSE]))[1]`

number of unique industry-occupation-area groups, `r dim(unique(merge_table[,c("NAICS", "OCC_CODE", "AREA_CODE"), with = FALSE]))[1]`

number of unique year-occupation-area groups, `r dim(unique(merge_table[,c("YEAR", "OCC_CODE", "AREA_CODE"), with = FALSE]))[1]`

## Four Variable Pairs
number of unique industry-year-occupation-area groups, `r dim(unique(merge_table[,c("NAICS", "YEAR", "OCC_CODE", "AREA_CODE"), with = FALSE]))[1]`

# OEWS Data
```{r OEWS_precode, echo = FALSE, results='asis', warning=FALSE, include = FALSE, message = FALSE}
OE_default <- c("YEAR", "NAICS", "OCC_CODE", "AREA_CODE")
OE_desc <- read.xlsx("Merge/OEWS_desc.xlsx", 1, rowIndex = 9:32, colIndex = 1:2)
colnames(OE_desc) <- OE_desc[1,]
OE_desc <- data.table(OE_desc[-1,])
OE_select <- gsub(" ", "_", toupper(OE_desc[12:dim(OE_desc)[1],]$Field))
OE <- fread("Merge/OEWS_agg.csv", select = c(OE_default, OE_select))
merge_table_OE <- OE[merge_table, on = c(OE_default)]
merge_table_OE[,(OE_select):= lapply(.SD, as.numeric), .SDcols = OE_select]
```

```{r OEWS_summary, echo = FALSE, results='asis'}
for (col in OE_select) {
  print(summaryStats(c(merge_table_OE %>% select(col))[[1]], 
               OE_desc[Field == tolower(ifelse(col == "LOC_QUOTIENT", "LOC QUOTIENT", col))]$`Field Description`, col))
  cat("\n")
}
```
# Labor Productivity Data
```{r LP_precode, echo = FALSE, results='asis', warning=FALSE, include = FALSE, message = FALSE}
#labor productivity
LP_default <- c("YEAR_IP", "NAICS_IP", "AREA_CODE_IP")
LP_desc <- fread("Merge/LP_desc.csv")
LP_select <- LP_desc$MEASURE_CODE_EXPANDED
IP <- fread("Merge/LP_agg.csv", select= c(LP_default,LP_select))
merge_table_IP <- IP[merge_table, on = c(LP_default)]
merge_table_IP[,(LP_select):= lapply(.SD, as.numeric), .SDcols = LP_select]
merge_table_IP[NAICS_IP == "", `NAICS_IP`:=NA]
```
## General Statistics
```{r LP_match, echo = FALSE, results='asis'}
perfect_match <- dim(merge_table_IP[NAICS == NAICS_IP & YEAR == YEAR_IP & AREA_CODE == AREA_CODE_IP])[1]
agg_match <- dim(merge_table_IP[LEVEL_NAICS_OE - LEVEL_NAICS_IP != 0])[1]
```
#### exact industry-year-area matches
number: `r perfect_match`

proportion: `r perfect_match/dim(merge_table_IP)[1]`


#### aggregated industry-year-area matches (industry was aggregated)
number: `r agg_match`

proportion: `r perfect_match/dim(merge_table_IP)[1]`

table of the number of digits that industries were aggregated by
```{r LP_match_2, echo = FALSE, results='asis'}
kable(table(merge_table_IP[LEVEL_NAICS_OE - LEVEL_NAICS_IP != 0, LEVEL_NAICS_OE - LEVEL_NAICS_IP]), col.names = c("digits of industry aggregation", "freq"))  %>%
  kable_classic_2(full_width = F)
```

#### missing industry-year-area matches
number: `r sum(is.na(merge_table_IP$NAICS_IP))`

proportion: `r mean(is.na(merge_table_IP$NAICS_IP))`

#### years covered
`r sort(unique(merge_table_IP$YEAR_IP))`


## Measure Statistics
```{r LP_summary, echo = FALSE, results='asis'}
for (col in LP_select) {
  if (col != "T30_1_IP") {
      print(summaryStats(c(merge_table_IP %>% select(col))[[1]], 
               LP_desc[MEASURE_CODE_EXPANDED == col, 
                             paste(MEASURE_TEXT, DURATION_TEXT)],
               col))
    cat("\n")
  }
  else {
    print("T30_1_IP - Value of production (Million $) Annual percent changes couldn't be summarized because of formatting issues, but is a derivative of T30_0_IP")
  }
}
```

# Employer Cost Data
````{r EC_precode, echo = FALSE, results='asis', warning=FALSE, include = FALSE, message = FALSE}
EC_default <- c("YEAR_CM", "INDUSTRY_CODE_CM", "OCCUPATION_CODE_CM", "AREA_CODE_CM")
EC_desc <- fread("Merge/EC_desc.csv")
EC_select <- EC_desc$estimate_datatype_code
EC <- fread("Merge/EC_agg.csv", select= c(EC_default,EC_select))
EC$INDUSTRY_CODE_CM <- as.numeric(EC$INDUSTRY_CODE_CM)
merge_table_EC <- EC[merge_table, on = c(EC_default)]
merge_table_EC[,(EC_select):= lapply(.SD, as.numeric), .SDcols = EC_select]
```
## General Statistics
```{r EC_match, echo = FALSE, results='asis', warning=FALSE, include = FALSE, message = FALSE}
perfect_match <- dim(merge_table_EC[(LEVEL_NAICS_OE - LEVEL_NAICS_CM) == 0 & 
                                      (LEVEL_OCC_OE - LEVEL_OCC_CM) == 0])[1]

agg_match_ind <- dim(merge_table_EC[(LEVEL_NAICS_OE - LEVEL_NAICS_CM) != 0 & 
                                      (LEVEL_OCC_OE - LEVEL_OCC_CM) == 0])[1]
```
#### exact industry-occupation-year-area matches
number, `r perfect_match`

proportion, `r perfect_match/dim(merge_table_EC)[1]`

#### aggregated industry-occuation-year-area matches (just industry was aggregated, occupation perfect match)
table of the number of industry digits that were aggregated

```{r EC_match2, echo = FALSE, results='asis', warning=FALSE, message = FALSE}
kable(table(merge_table_EC[(LEVEL_NAICS_OE - LEVEL_NAICS_CM) != 0 & 
                             (LEVEL_OCC_OE - LEVEL_OCC_CM) == 0, LEVEL_NAICS_OE - LEVEL_NAICS_CM]), col.names = c("Number of digits industries aggregated by", "freq")) %>%
  kable_classic_2(full_width = F)

agg_match_occ <- dim(merge_table_EC[(LEVEL_NAICS_OE - LEVEL_NAICS_CM) == 0 & 
                                      (LEVEL_OCC_OE - LEVEL_OCC_CM) != 0])[1]
```

#### aggregated industry-occuation-year-area matches (just occupation was aggregated, industry perfect match)
number, `r agg_match_occ`

proportion, `r agg_match_occ/dim(merge_table_EC)[1]`

table of the number of digits occupations were aggregated by

```{r EC_match3, echo = FALSE, results='asis', warning=FALSE, message = FALSE}
kable(table(merge_table_EC[(LEVEL_NAICS_OE - LEVEL_NAICS_CM) == 0 & 
                       (LEVEL_OCC_OE - LEVEL_OCC_CM) != 0, LEVEL_OCC_OE - LEVEL_OCC_CM]),
      col.names = c("Number of digits occupations were aggregated by", "freq"))  %>%
  kable_classic_2(full_width = F)
temp_match <- merge_table_EC[(LEVEL_NAICS_OE - LEVEL_NAICS_CM) != 0 & 
                               (LEVEL_OCC_OE - LEVEL_OCC_CM) != 0]
temp_match[, `occ_dif`:= LEVEL_OCC_OE - LEVEL_OCC_CM]
temp_match[, `ind_dif`:= LEVEL_NAICS_OE - LEVEL_NAICS_CM]
agg_match_occ_ind <- dim(temp_match)[1]
```

#### aggregated industry-occuation-year-area matches (occupation & industry were both aggregated)
number, `r agg_match_occ_ind`

proportion, `r agg_match_occ_ind/dim(merge_table_EC)[1]`

table of the number of digits occupation-industry pairs were aggregated by
```{r EC_match4, echo = FALSE, results='asis', warning=FALSE, message = FALSE}
kable(table(temp_match[,c("occ_dif", "ind_dif"), with = FALSE]), row.names = c(1, 2, 3, 4))  %>%
  kable_classic_2(full_width = F)
```

#### missing industry-occupation-year-area matches
number, `r sum(is.na(merge_table_EC$INDUSTRY_CODE_CM))`

proportion, `r mean(is.na(merge_table_EC$INDUSTRY_CODE_CM))`


#### Years covered
`r sort(unique(merge_table_EC$YEAR_CM))`


## Measure Statistics
```{r EC_summary, echo = FALSE, results='asis', warning=FALSE, message = FALSE}
for (col in EC_select) {
  print(summaryStats(c(merge_table_EC %>% select(col))[[1]], 
               EC_desc[estimate_datatype_code == col, 
                       paste(estimate_text, datatype_text)],
               col))
  cat("\n")
}
```

# Wage Model Data
```{r WM_precode, echo = FALSE, results='asis', warning=FALSE, message = FALSE}
WM_default <- c("YEAR_WM", "OCCUPATION_CODE_WM", "AREA_CODE_WM")
WM_desc <- fread("Merge/WM_desc.csv", sep="^")
WM_select <- WM_desc$MEASURE
WM <- fread("Merge/WM_agg.csv", select= c(WM_default,WM_select))
merge_table_WM <- WM[merge_table, on = c(WM_default)]
merge_table_WM[,(WM_select):= lapply(.SD, as.numeric), .SDcols = WM_select]

perfect_match <- dim(merge_table_WM[LEVEL_OCC_OE - LEVEL_OCC_WM == 0])[1]
agg_match_occ <- dim(merge_table_WM[LEVEL_OCC_OE - LEVEL_OCC_WM != 0])[1]
```

## General Statistics
#### exact occupation-year-area matches
number, `r perfect_match`

proportion, `r perfect_match/dim(merge_table_WM)[1]`


#### aggregated occuation-year-area matches
number, `r agg_match_occ`

proportion, `r agg_match_occ/dim(merge_table_WM)[1]`

table of the number of digits occupations were aggregated by

```{r WM_match2, echo = FALSE, results='asis', warning=FALSE, message = FALSE}
kable(table(merge_table_WM[LEVEL_OCC_OE - LEVEL_OCC_WM != 0, LEVEL_OCC_OE - LEVEL_OCC_WM]),
      col.names = c("# of digits of occupation aggregation", "freq")) %>%
  kable_classic_2(full_width = F)
```

#### missing occupation-year-area matches
number, `r sum(is.na(merge_table_WM$OCCUPATION_CODE_WM))`

proportion, `r mean(is.na(merge_table_WM$OCCUPATION_CODE_WM))`

#### Years covered
`r sort(unique(merge_table_WM$YEAR_WM))`

## Measure Statistics
```{r WM_summary, echo = FALSE, results='asis', warning=FALSE, message = FALSE}
for (col in WM_select) {
  print(summaryStats(c(merge_table_WM %>% select(col))[[1]], 
               WM_desc[MEASURE == col, MEASURE],
               col))
}
```


